Paper

### 2021-12-16 Paper.Rmd

Background:
The paper "Novel approach for Monte Carlo simulation of the new COVID-19 spread dynamics aim's to investigate the transmission and incubation random processes of COVID-19. The goal of the study is to generate epidemiological data based on the natural mechanism's of transmission of the virus assuming random interactions of a large-finite number of individuals in very short distance ranges. A few important assumptions that were made were in how the transmission of the virus occurs when the physical distance between individuals decreases to where the probability of transmission converges to one, asymptotically. Their model defined the steps of movement in a defined space as a randomly assigned process. They approached this through the SIR model; A model defining Susceptible, Infected, Recovered individuals. All change in the SIR model is based on time, as the function iterates through a set number of days. They started with a 0.5% infection rate and simulation daily infection counts over a range of time. In order to capture many factors in the transmission including the relationship between infection probability between physical distance, effective and ineffective mask wearing, and environmental conditions they distributed these factors Exponential(rate = lambda). Initial recovery and incubation times were distributed Gamma(alpha,beta).

@@

```
In the previous three simulations we assumed that no one was vaccinated, but we also wanted to look at how a vaccine would effect the spread of Covid-19. In these simulations we looked at 200 people with the same dimension parameters as before, but with only one person initially infected. We increased the infection distance since 200 people in a 1.3 square mile area is an even lower population density than before. All the parameters were held the same other than the vaccination rate for these simulations. The vaccine rates of the four examples were: 0, .5, .8, and .95. In the plot that shows the number of infected for each vaccination rate, black is 0 vaccinated, red is .5, blue is .8, and green is .95. As we would expect, when no one is vaccinated, the number of infected people peaks and falls much faster than when some people are vaccinated. It only took around 15 days to peak and nearly everyone was recovered by day 30. When no one was vaccinated it took much less time to get sick, but people also were recovered faster because everyone already had the virus. This is the idea of herd immunity that is often talked about, but it poses a massive problem with the hospitals. If everyone gets sick in a short period of time, hospitals can get overwhelmed and the death tolls can skyrocket due to lack of equipment and staff. Vaccinations aren't perfect, but they can help flatten the curve so the infection numbers don't peak as fast and also reduce the severity of the symptoms so less people need to be hospitalized. It is still possible to contract Covid-19 while vaccinated, but it is much less likely. As you can see in the last plot, as the vaccination rate increased, the longer it took for the cases to peak. The peak was also much lower than when no one was vaccinated and it took longer for the cases to completely disappear since people were being infected slower. This shows that a higher vaccination rate helps slow down the spread of Covid-19 even if it can't completely prevent it.

>>>>

Background:
The paper "Novel approach for Monte Carlo simulation of the new COVID-19 spread dynamics aim's to investigate the transmission and incubation random processes of COVID-19. The goal of the study is to generate epidemiological data based on the natural mechanism's of transmission of the virus assuming random interactions of a large-finite number of individuals in very short distance ranges. A few important assumptions that were made were in how the transmission of the virus occurs when the physical distance between individuals decreases to where the probability of transmission converges to one, asymptotically. Their model defined the steps of movement in a defined space as a randomly assigned process. They approached this through the SIR model; A model defining Susceptible, Infected, Recovered individuals. All change in the SIR model is based on time, as the function iterates through a set number of days. They started with a 0.5% infection rate and simulation daily infection counts over a range of time. In order to capture many factors in the transmission including the relationship between infection probability between physical distance, effective and ineffective mask wearing, and environmental conditions they distributed these factors Exponential(rate = lambda). Initial recovery and incubation times were distributed Gamma(alpha,beta). (Source 1)

@@

In the previous three simulations we assumed that no one was vaccinated, but we also wanted to look at how a vaccine would effect the spread of Covid-19. In these simulations we looked at 200 people with the same dimension parameters as before, but with only one person initially infected. We increased the infection distance since 200 people in a 1.3 square mile area is an even lower population density than before. All the parameters were held the same other than the vaccination rate for these simulations. The vaccine rates of the four examples were: 0, .5, .8, and .95. In the plot that shows the number of infected for each vaccination rate, black is 0 vaccinated, red is .5, blue is .8, and green is .95. As we would expect, when no one is vaccinated, the number of infected people peaks and falls much faster than when some people are vaccinated. It only took around 15 days to peak and nearly everyone was recovered by day 30. When no one was vaccinated it took much less time to get sick, but people also were recovered faster because everyone already had the virus. This is the idea of herd immunity that is often talked about, but it poses a massive problem with the hospitals. If everyone gets sick in a short period of time, hospitals can get overwhelmed and the death tolls can skyrocket due to lack of equipment and staff. Vaccinations aren't perfect, but they can help flatten the curve so the infection numbers don't peak as fast and also reduce the severity of the symptoms so less people need to be hospitalized. It is still possible to contract Covid-19 while vaccinated, but it is much less likely. As you can see in the last plot, as the vaccination rate increased, the longer it took for the cases to peak. The peak was also much lower than when no one was vaccinated and it took longer for the cases to completely disappear since people were being infected slower. This shows that a higher vaccination rate helps slow down the spread of Covid-19 even if it can't completely prevent it.

Bibliography:

[1] Maltezos, S., & Georgakopoulou, A. (2021). Novel approach for Monte Carlo simulation of the new COVID-19 spread dynamics. Infection, Genetics and Evolution, 92, 104896. https://doi.org/10.1016/j.meegid.2021.104896

[2] Lauer, S. A., Grantz, K. H., Bi, Q., Jones, F. K., Zheng, Q., Meredith, H. R., Azman, A. S., Reich, N. G., & Lessler, J. (2020). The Incubation Period of Coronavirus Disease 2019 (COVID-19) From Publicly Reported Confirmed Cases: Estimation and Application. Annals of internal medicine, 172(9), 577–582. https://doi.org/10.7326/M20-0504

[3] H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.

[4] H. Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686

[5] Angelo Canty and Brian Ripley (2021). boot: Bootstrap R (S-Plus) Functions. R package version 1.3-28., Davison, A. C. & Hinkley, D. V. (1997) Bootstrap Methods and Their Applications. Cambridge University Press, Cambridge. ISBN 0-521-57391-2
